Litrepl: Literate Paper Processor Promoting Transparency More Than Reproducibility

### 2024-12-29 paper/paper.tex
Update paper

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{svg}
\usepackage{comment}

@@

@@

\begin{comment}
\begin{sh}
echo '\begin{verbatim}'
echo '$ cat input.tex'
cat input.tex
echo '\end{verbatim}'
\end{sh}
\end{comment}
%lresult
\begin{verbatim}
$ cat input.tex
\begin{python}
import sys
print(f"Hello, {sys.platform}!")
\end{python}
\begin{result}
\end{result}
\end{verbatim}
%lnoresult

@@

\begin{comment}
\begin{sh}
echo '\begin{verbatim}'
echo '$ cat input.tex | litrepl'
echo "sys.platform='linux'" | litrepl repl python >/dev/null
cat input.tex | litrepl
echo '\end{verbatim}'
\end{sh}
\end{comment}
%lresult
\begin{verbatim}
$ cat input.tex | litrepl
\begin{python}
import sys
print(f"Hello, {sys.platform}!")
\end{python}
\begin{result}
Hello, linux!
\end{result}
\end{verbatim}
%lnoresult

Now we can see the expected greeting of the author's operating system platform.
The side-effect of this execution is the started session of the python
interpreter which is now running in the background. We can modify its state by
adding more section to the document and executing them selectively or e.g. by
accessing \verb|litrepl repl python| terminal. So for example, setting
\verb|sys.platform| to another value and re-evaluating the document would yield
a different greeting.

\begin{figure*}[t]
  \centering
  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
  \caption{Caption describing the image shown in the figure.}
  \label{fig:resource-allocation}
\end{figure*}

@@

@@

When an evaluation request is received, Litrepl calculates the name of the
anticipated response storage file and forwards the request to the interpreter.
It concurrently initiates a response reader process. The response storage is
secured using soft POSIX file locks, enabling the main process to ascertain when
the response is ready. In instances where the response exceeds the configured
time to completion, Litrepl provides a partial result. This partial result can
be updated with complete details in subsequent calls. The format for a partial
result section appears as follows:
% Displaying a verbatim box above two columns in a two-column paper

\begin{figure*}[t]
  \centering
  \begin{minipage}{\textwidth}
    \centering
    %lresult
    \begin{verbatim}
    \begin{result}
    ... some output ...
    [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
    \end{result}
    \end{verbatim}
    %lnoresult
  \end{minipage}
  \caption{Example output in a verbatim format displayed above two columns.}
  \label{fig:verbatim_example}
\end{figure*}

% The figure* environment is used to span two columns, while a minipage is
% used to contain the verbatim listing. This code creates a box above the
% two-column layout with a caption explaining the contents.

@@

Details and discussions about the second subsection.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{svg}
\usepackage{comment}
\usepackage{float}

@@

\begin{figure*}[!hbt]
  \centering
  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
  \caption{Caption describing the image shown in the figure.}
  \label{fig:resource-allocation}
\end{figure*}

@@

\begin{comment}
\begin{sh}
echo '\begin{verbatim}'
echo '$ cat input.tex'
cat input.tex
echo '\end{verbatim}'
\end{sh}
\end{comment}
%lresult
\begin{verbatim}
$ cat input.tex
\begin{python}
import sys
print(f"I use {sys.platform} btw!")
\end{python}
\begin{result}
\end{result}
\end{verbatim}
%lnoresult

@@

\begin{comment}
\begin{sh}
echo '\begin{verbatim}'
echo '$ cat input.tex | litrepl'
echo "sys.platform='linux'" | litrepl repl python >/dev/null
cat input.tex | litrepl
echo '\end{verbatim}'
\end{sh}
\end{comment}
%lresult
\begin{verbatim}
$ cat input.tex | litrepl
\begin{python}
import sys
print(f"I use {sys.platform} btw!")
\end{python}
\begin{result}
I use linux btw!
\end{result}
\end{verbatim}
%lnoresult

Now we can see the expected statement about the author's operating system.  The
side-effect of this execution is the started session of the python interpreter
which is now running in the background. We can modify its state by adding more
section to the document and executing them selectively or e.g. by accessing
\verb|litrepl repl python| terminal. So for example, setting \verb|sys.platform|
to another value and re-evaluating the document would yield a different
statement.

@@

\begin{figure*}[hbt!]
  \centering
  %lresult
  \begin{verbatim}
  \begin{result}
  ... some output ...
  [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
  \end{result}
  \end{verbatim}
  %lnoresult
  \caption{Example output in a verbatim format displayed above two columns.}
  \label{fig:verbatim_example}
\end{figure*}

@@

When an evaluation request is received, Litrepl calculates the name of the
anticipated response storage file and forwards the request to the interpreter.
It concurrently initiates a response reader process. The response storage is
secured using soft POSIX file locks, enabling the main process to ascertain when
the response is ready. In instances where the response exceeds the configured
time to completion, Litrepl provides a partial result. This partial result can
be updated with complete details in subsequent calls. The format for a partial
result section appears as follows:

@@

\begin{figure*}[hbt!]
  \centering
  \begin{verbatim}
  document ::= (code_section | result_section | ignore_section | text)*
  code_section ::= code_bmarker code_section_text code_emarker
  result_section ::= result_bmarker result_section_text result_emarker
  ignore_section ::= ignore_bmarker ignore_section_text ignore_emarker
  \end{verbatim}
  \caption{Example output in a verbatim format displayed above two columns.}
  \label{fig:document_representation}
\end{figure*}

To efficiently detect sections within a document, Litrepl employs simplified
versions of each document markup grammar, which it defines and parses. In this
approach, documents are abstracted as a straightforward sequence comprising
code, result, and text sections. Additionally, Litrepl identifies ignore blocks,
which serve as block comments. Any sections encapsulated within these ignore
blocks are excluded from parsing and evaluation. The structure of any document,
therefore, can be articulated using the following informal grammar:

The specific markers that denote each block type are determined by the document
format and configuration settings. For example, in a LaTeX document containing
Python code, Litrepl recognizes markers such as \verb|\begin|\verb|{python}| or
\verb|% |\verb|python| as code section begin markers, and
\verb|\end|\verb|{python}| or \verb|% |\verb|nopython| as end markers. These
latter variants are transparent to LaTeX, facilitating the inclusion of LaTeX
markup seamlessly.

This parsing strategy allows Litrepl to navigate and process documents with
precision, ensuring that relevant sections are evaluated while non-relevant
parts are effectively ignored. By integrating document format-specific markers,
Litrepl maintains flexibility and adaptability in diverse document scenarios.
